Estimating Uncertainty for Tipping Points with Deep Learning

### 2022-05-04 manuscript/manuscript.Rmd
use hive partition structure, example plotting

```{r fig.cap="Forecast scenarios. A. a Hopf bifurcation: a stable node develops into a limit cycle. B. The saddle-node bifurcation."}
hopf <- read_csv("../greta/data/hopf_increasing.csv.gz")
panelA <- hopf |> filter(type %in% c("historical", "true"), i < 10) |>
  rename(species = variable) |> 
  mutate(species = forcats::fct_recode(species, host = "H", parasitoid = "P")) |>
  ggplot(aes(t, value, group=interaction(species, i, type), col=species)) + 
  geom_line(show.legend = FALSE, alpha = 0.7) + theme_bw()+  scale_color_grey() +
  ggtitle("A. Hopf bifurcation")

sn <- read_csv("../greta/data/saddlenode.csv.gz")
panelB <- sn |> filter(type == "historical", i == 1) |>
  bind_rows(filter(sn, type=="true")|>mutate(type="predicted")) |>
  ggplot(aes(t, value, col = type, group=interaction(i,type), lty=type)) +
  geom_path(alpha=0.5, show.legend = FALSE) +
  scale_color_grey(end=0.6) + theme_bw() + 
  ggtitle("B. Saddle-node bifurcation")

stoch <- read_csv("../greta/data/stochastic.csv.gz")
panelC <- stoch |> filter(type == "historical", i == 1) |>
  bind_rows(filter(stoch, type=="true", i < 50) |> mutate(type="predicted")) |>
  ggplot(aes(t, value, col = type, group=interaction(i,type), lty=type)) +
  geom_path(alpha=0.5) + scale_color_grey(end=0.6) + theme_bw() + 
  ggtitle("C. Stochastic transition")

@@

```{r}
panelB <- sn |> filter(type == "historical", i == 1) |>
  bind_rows(filter(sn, type=="true")|>mutate(type="predicted")) |>
  ggplot(aes(t, value, col = type, group=interaction(i,type), lty=type)) +
  geom_path(alpha=0.5, show.legend = FALSE) +
  scale_color_grey(end=0.6) + theme_bw() + 
  ggtitle("B. Saddle-node bifurcation")

panelC <- stoch |> filter(type == "historical", i == 1) |>
  bind_rows(filter(stoch, type=="true", i < 50) |> mutate(type="predicted")) |>
  ggplot(aes(t, value, col = type, group=interaction(i,type), lty=type)) +
  geom_path(alpha=0.5) + scale_color_grey(end=0.6) + theme_bw() + 
  ggtitle("C. Stochastic transition")

@@

```{r}
mcmc_scores <- bind_rows(
  read_csv("../greta/data/scores_saddlenode.csv.gz") |> mutate(scenario="saddlenode"),
read_csv("../greta/data/scores_stochastic.csv.gz"),
read_csv("../greta/data/scores_hopf.csv.gz") |> 
  filter(scenario == "hopf_increasing") |>
  mutate(scenario = forcats::fct_recode(scenario, hopf = "hopf_increasing"),
         reps = 1)
) |> rename(method=model)
mcmc_scores |> 
  pivot_longer(c(logs, crps), names_to = "metric", values_to = "score") |> 
  ggplot(aes(method, score, col=variable)) + geom_boxplot() + 
  facet_grid(metric~scenario, scales="free_y")
```

>>>>

```{r fig.cap="Forecast scenarios. A. a Hopf bifurcation: a stable node develops into a limit cycle. B. The saddle-node bifurcation."}
hopf <- read_csv("../greta/forecasts/hopf_increasing.csv.gz")
panelA <- hopf |> filter(type %in% c("historical", "true"), i < 10) |>
  rename(species = variable) |> 
  mutate(species = forcats::fct_recode(species, host = "H", parasitoid = "P")) |>
  ggplot(aes(t, value, group=interaction(species, i, type), col=species)) + 
  geom_line(show.legend = FALSE, alpha = 0.7) + theme_bw()+  scale_color_grey() +
  ggtitle("A. Hopf bifurcation")

sn <- read_csv("../greta/forecasts/saddlenode.csv.gz")
panelB <- sn |> filter(type == "historical", i == 1) |>
  bind_rows(filter(sn, type=="true")|>mutate(type="predicted")) |>
  ggplot(aes(t, value, col = type, group=interaction(i,type), lty=type)) +
  geom_path(alpha=0.5, show.legend = FALSE) +
  scale_color_grey(end=0.6) + theme_bw() + 
  ggtitle("B. Saddle-node bifurcation")

stoch <- read_csv("../greta/forecasts/stochastic.csv.gz")
panelC <- stoch |> filter(type == "historical", i == 1) |>
  bind_rows(filter(stoch, type=="true", i < 50) |> mutate(type="predicted")) |>
  ggplot(aes(t, value, col = type, group=interaction(i,type), lty=type)) +
  geom_path(alpha=0.5) + scale_color_grey(end=0.6) + theme_bw() + 
  ggtitle("C. Stochastic transition")

@@

```{r}
panelB <- sn |>
  group_by(t,type,variable) |> 
  summarise(mean = mean(value), sd = sd(value), .groups = "drop") |> 
  rename(species = variable) |>
  ggplot(aes(t, col=type)) + 
    geom_ribbon(aes(ymin = mean-2*sd, ymax = mean+2*sd, fill=type), alpha=0.5) +
    geom_line(aes(y=mean)) +
    #geom_vline(aes(xintercept = 100)) + 
    facet_wrap(~species, ncol=1) +
    theme_bw()

panelC <- stoch |>
  group_by(t,type,variable) |> 
  summarise(mean = mean(value), sd = sd(value), .groups = "drop") |> 
  rename(species = variable) |>
  ggplot(aes(t, col=type)) + 
    geom_ribbon(aes(ymin = pmax(mean-2*sd,0), ymax = mean+2*sd, fill=type), alpha=0.5) +
    geom_line(aes(y=mean)) +
    #geom_vline(aes(xintercept = 100)) + 
    facet_wrap(~species, ncol=1) +
    theme_bw() + 
  ggtitle("C. Stochastic transition")

@@

```{r}
mcmc_scores <- bind_rows(
  read_csv("../greta/forecasts/scores_saddlenode.csv.gz") |> mutate(scenario="saddlenode"),
read_csv("../greta/forecasts/scores_stochastic.csv.gz"),
read_csv("../greta/forecasts/scores_hopf.csv.gz") |> 
  filter(scenario == "hopf_increasing") |>
  mutate(scenario = forcats::fct_recode(scenario, hopf = "hopf_increasing"),
         reps = 1)
) |> rename(method=model)
mcmc_scores |> 
  pivot_longer(c(logs, crps), names_to = "metric", values_to = "score") |> 
  ggplot(aes(method, score, col=variable)) + geom_boxplot() + 
  facet_grid(metric~scenario, scales="free_y")
```
